---
title: "New York Cases and Deaths and COVID-19"
output:
  html_document:
    df_print: paged
---

# Associating the C-19 Community Risk Score with deaths in NYC
- Chirag Patel (chirag@xy.ai)
- 05/20/20
- As of 5/19/2020, New York City released a zipcode-level case and death count due to COVID-19. 
- Is COVID-19 Community Risk Score associated with COVID-19 related deaths in NYC?
- modified 09/26/20 with new September death rates
- see: https://github.com/nychealth/coronavirus-data

```{r load census and death data}
library(tidyverse)
library(ggrepel)
library(MASS)
library(stargazer)
library(sjPlot)
library(cowplot)
```

# Load NYC data
```{r}
## MAY 2020
nyc_deaths_may <- read_csv('./external_data/data-by-modzcta_052120.csv') #or: ead_csv(url('https://raw.githubusercontent.com/nychealth/coronavirus-data/master/data-by-modzcta.csv')) # current for 5/20/20
nyc_deaths_may <- nyc_deaths_may %>% mutate(TOTAL_COVID_TESTS = NA, date_download='05/20/20')

## SEPT 2020
nyc_deaths_sept <- read_csv('./external_data/data-by-modzcta_092620.csv') 
nyc_deaths_sept <- nyc_deaths_sept %>% mutate(date_download = '09/26/20')

nyc_deaths <- rbind(nyc_deaths_sept, nyc_deaths_may)
zcta_to_census <- read_csv(url('http://www2.census.gov/geo/docs/maps-data/data/rel/zcta_tract_rel_10.txt?#')) # census walk file from tract to ZCTA
nyc_deaths$MODIFIED_ZCTA <- as.character(nyc_deaths$MODIFIED_ZCTA)
nyc_census <- nyc_deaths %>% left_join(zcta_to_census, by=c('MODIFIED_ZCTA'='ZCTA5'))


```

```{r load raw and risk score data}
## load the raw data
source('data_prep_risk_score.R')
risk_score <- read_rds('./scores/fh_acs_covid_comm_score.rds')

# combine with SES variables
risk_score_acs <- risk_score %>% left_join(fh_acs, by=c('geoid'='geoid'))
nyc_census$GEOID <- as.character(nyc_census$GEOID)
fh_acs_zip <- fh_acs %>% right_join(nyc_census %>% dplyr::select(c(MODIFIED_ZCTA, GEOID)), by=c('geoid'='GEOID'))
risk_score_nyc <- risk_score_acs %>% right_join(nyc_census %>% dplyr::select(c(MODIFIED_ZCTA, GEOID, ZPOPPCT)), by=c('geoid'='GEOID'))


risk_score_nyc <- risk_score_nyc %>% group_by(MODIFIED_ZCTA) %>% summarize(risk_score=weighted.mean(risk_score_both_pc, ZPOPPCT),
                                                                          median_income = weighted.mean(b19013001,ZPOPPCT),
                                                                          median_home_value = weighted.mean(b25077001, ZPOPPCT),
                                                                          no_health_insurance_pct=sum(b27001005 + b27001008 + b27001011 + b27001014 + b27001017 + b27001020 + b27001023 + b27001026 + b27001029 + b27001033 + b27001036 + b27001039 + b27001042 + b27001045 + b27001048 + b27001051 + b27001054 + b27001057)/sum(b27001001),
                                                                          african_american_pct = sum(c02003004) / sum(c02003001),
                                                                          mexican_pct = sum(b03001004) / sum(b03001001),
                                                                          hispanic_pct = sum(b03001003) / sum(b03001001),
                                                                          asian_pct = sum(c02003006) / sum(c02003001),
                                                                          white_pct = sum(c02003003) / sum(c02003001),
                                                                          unemployment_pct = sum(b23025005 ) / sum(b23025001),
                                                                          less_than_high_school_pct = sum(b15003002 + b15003003 + b15003004 + b15003005 + b15003006 + b15003007 + b15003008 + b15003009 + b15003010 + b15003011 + b15003012 + b15003013 + b15003014 + b15003015 + b15003016) / sum(b15003001),
                                                                          college_pct = sum(b15003021 + b15003022 + b15003023 + b15003024 + b15003025) / sum(b15003001),
                                                                          at_below_poverty_pct = sum(b17001003 + b17001017 ) / sum(b17001001),
                                                                          more_than_one_occupant_per_room_pct = sum(b25014005 + b25014006 + b25014007 + b25014011 + b25014012 + b25014013) / sum(b25014001)
)
                                                                           
risk_score_nyc <- risk_score_nyc %>% right_join(nyc_deaths)
```

# COVID-19 Risk Score vs. Case and Death Rate (zipcode level)

```{r plot cases and deaths}


p  <- ggplot(risk_score_nyc, aes(COVID_CASE_RATE, COVID_DEATH_RATE, color=date_download))
p <- p + geom_point()
p

p  <- ggplot(risk_score_nyc, aes(risk_score, COVID_CASE_RATE, color=date_download))
p <- p + geom_point()
p

p  <- ggplot(risk_score_nyc, aes(risk_score, COVID_DEATH_RATE, color=date_download))
p <- p + geom_point()
p <- p + geom_text_repel(data=risk_score_nyc %>% filter(risk_score > 45), aes(risk_score, COVID_DEATH_RATE, label=MODIFIED_ZCTA))
p



```

# Write out the risk score per death rate for plotting
```{r}
write_csv(risk_score_nyc %>% dplyr::select(risk_score, COVID_DEATH_RATE, MODIFIED_ZCTA,NEIGHBORHOOD_NAME, NEIGHBORHOOD_NAME) %>%  arrange(desc(risk_score)), path = './scores/new_york_risk_score_death_092720.csv')

```

# Model the death rate

```{r results='asis', message = FALSE, fig.width=12, fig.height=6}

mod <- lm(COVID_DEATH_RATE ~ risk_score + POP_DENOMINATOR, risk_score_nyc %>% filter(date_download == '05/20/20'))
mod_2 <- lm(COVID_DEATH_RATE ~ risk_score + POP_DENOMINATOR + COVID_CASE_RATE, risk_score_nyc %>% filter(date_download == '05/20/20'))
#mod_3 <- glm( COVID_DEATH_COUNT ~ risk_score + offset(log(POP_DENOMINATOR)), data=risk_score_nyc, family="poisson")

mod_5 <- glm.nb( COVID_DEATH_COUNT ~ risk_score + median_income +  less_than_high_school_pct +college_pct + african_american_pct + mexican_pct + hispanic_pct + asian_pct  + at_below_poverty_pct + more_than_one_occupant_per_room_pct + no_health_insurance_pct + unemployment_pct + COVID_CASE_COUNT + offset(log(POP_DENOMINATOR)), data=risk_score_nyc %>% filter(date_download == '05/20/20'))

mod_5_sept <- glm.nb( COVID_DEATH_COUNT ~ risk_score + median_income +  less_than_high_school_pct +college_pct + african_american_pct + mexican_pct + hispanic_pct + asian_pct  + at_below_poverty_pct + more_than_one_occupant_per_room_pct + no_health_insurance_pct + unemployment_pct + COVID_CASE_COUNT + offset(log(POP_DENOMINATOR)), data=risk_score_nyc %>% filter(date_download == '09/26/20'))


mod_6 <- lm(COVID_DEATH_RATE ~ I(risk_score/10) + median_income +  less_than_high_school_pct + college_pct + african_american_pct + mexican_pct + hispanic_pct + asian_pct + at_below_poverty_pct + more_than_one_occupant_per_room_pct + no_health_insurance_pct + unemployment_pct + POP_DENOMINATOR, data=risk_score_nyc %>% filter(date_download == '05/20/20'))

mod_7 <- lm(COVID_DEATH_RATE ~ I(risk_score/10) + median_income +  less_than_high_school_pct +college_pct + african_american_pct + mexican_pct + hispanic_pct + asian_pct + at_below_poverty_pct + more_than_one_occupant_per_room_pct + no_health_insurance_pct + unemployment_pct + COVID_CASE_COUNT + POP_DENOMINATOR, data=risk_score_nyc %>% filter(date_download == '05/20/20'))


mod_8 <- lm(COVID_DEATH_RATE ~  median_income +  less_than_high_school_pct +college_pct + african_american_pct + mexican_pct + hispanic_pct + asian_pct + white_pct + at_below_poverty_pct + more_than_one_occupant_per_room_pct + no_health_insurance_pct + unemployment_pct + COVID_CASE_COUNT + POP_DENOMINATOR, data=risk_score_nyc %>% filter(date_download == '05/20/20'))



mod_5b <- glm.nb( COVID_DEATH_COUNT ~ I(risk_score > 40) + median_income +  less_than_high_school_pct +college_pct + african_american_pct + mexican_pct + hispanic_pct + asian_pct  + at_below_poverty_pct + more_than_one_occupant_per_room_pct + no_health_insurance_pct + unemployment_pct + COVID_CASE_COUNT + offset(log(POP_DENOMINATOR)), data=risk_score_nyc %>% filter(date_download == '05/20/20'))

may_model_plot <- plot_model(mod_5, type="std2", axis.lim = c(.25, 3))
sept_model_plot <- plot_model(mod_5_sept, type="std2", axis.lim = c(.25, 3))
plot_grid(may_model_plot,sept_model_plot, labels = c('May 2020', 'Sept 2020'))

```