---
output:
  html_document:
    df_print: paged
---

#Covid-19 Community Score

- last updated: 03/30/2020
- chirag@xy.ai


```{r}
library(ggpubr)
library(tidyverse)
library(readr)
library(gplots)
library(ggrepel)
library(ggthemes)
library(DT)
library(scales)
```

```{r load and prepare data}
fh_acs <- read_rds('./fh_acs.rds')
fh_names <- read_rds('./fh_orig_tract.rds')
fh_acs <- fh_acs %>% left_join(fh_names, by = c('geoid'='fips_tract'))
place_vars <- names(fh_names)
fh_acs <- fh_acs %>%  mutate(county_code =str_sub(geoid, 1, 5))
fh_acs <- fh_acs %>%  mutate(state_code =str_sub(geoid, 1, 2))

fh_vars_all <- names(fh_acs)[grep('Crude', names(fh_acs))]
fh_acs[, fh_vars_all] <- fh_acs[, fh_vars_all]/100 # get it to fraction units

fh_vars <- names(fh_acs)[grep('CrudePrev', names(fh_acs))]
health_prefix <- strsplit(fh_vars, '_') %>% map(function(arr) {arr[[1]]}) %>% unlist()
health_prefix <- setdiff(health_prefix, 'COLON')
fh_vars <- c(setdiff(fh_vars, 'COLON_SCREEN_CrudePrev'), 'COLON_SCREEN_CrudePrev')
health_prefix <- c(health_prefix, 'COLON_SCREEN')
fh_95hi_vars <- sprintf('%s_Crude95CI_high', health_prefix)
fh_se_vars <- sprintf('%s_SE', health_prefix)
fh_sd_vars <- sprintf('%s_SD', health_prefix)
fh_acs <- fh_acs %>% mutate(total_population = total_males + total_females)
for(index in 1:length(fh_vars)) {
  fh_acs[,fh_se_vars[index]] <- (fh_acs[,fh_95hi_vars[index]] - fh_acs[, fh_vars[index]])/1.96
  fh_acs[,fh_sd_vars[index]] <- fh_acs[, fh_se_vars[index]] * sqrt(fh_acs[,'total_population'])
}

fh_behavior <- c('BINGE', 'SLEEP', 'CSMOKING')
fh_disease <- c('ARTHRITIS', 'CANCER', 'CASTHMA', 'CHD', 'COPD', 'DIABETES', 'KIDNEY', 'TEETHLOST', 'STROKE')
fh_risk <- c('OBESITY', 'BPHIGH', 'LPA', 'HIGHCHOL')
fh_access <- c('ACCESS2', 'BPMED', 'CHECKUP', 'CHOLSCREEN', 'COLON_SCREEN', 'COREM', 'COREW', 'MAMMOUSE', 'MHLTH', 'PHLTH', 'PAPTEST')

fh_behavior <- fh_vars[fh_behavior %>% map(grep, fh_vars) %>% unlist()]
fh_disease <- fh_vars[fh_disease %>% map(grep, fh_vars) %>% unlist()]
fh_risk <- fh_vars[fh_risk %>% map(grep, fh_vars) %>% unlist()]
fh_access <- fh_vars[fh_access %>% map(grep, fh_vars) %>% unlist()]

```

```{r hardcode column names}
ses_vars <- c('gini_index')
ses_vars <- c(ses_vars, 'median_income')
ses_vars <- c(ses_vars, 'median_home_value')
ses_vars <- c(ses_vars, 'male_over_65_pct', 'female_over_65_pct')
ses_vars <- c(ses_vars, 'white_pct', 'african_american_pct', 'american_indian_pct', 'asian_pct', 'hawaiian_pacific_islander_pct', 'other_race_pct', 'two_or_more_race_pct')
ses_vars <- c(ses_vars, 'hispanic_pct', 'mexican_pct', 'puerto_rican_pct', 'cuban_pct', 'dominican_pct', 'central_american_pct', 'south_american_pct', 'other_hispanic_pct')
ses_vars <- c(ses_vars, 'less_than_high_school_pct', 'college_pct')
ses_vars <- c(ses_vars, 'at_below_poverty_pct', 'at_below_poverty_over_65_pct')
ses_vars <- c(ses_vars, 'unemployment_pct', 'non_employment_pct')
ses_vars <- c(ses_vars, 'more_than_one_occupant_per_room_pct', 'more_than_one_occupant_per_room_renter_pct')
ses_vars <- c(ses_vars, 'no_health_insurance_pct', 'no_health_insurance_over_65_pct')


established_disease <- c('CANCER_CrudePrev', 'ARTHRITIS_CrudePrev',  'STROKE_CrudePrev', 'CASTHMA_CrudePrev','COPD_CrudePrev', 'CHD_CrudePrev', 'DIABETES_CrudePrev', 'KIDNEY_CrudePrev', 'BPMED_CrudePrev')
established_risk <- c( 'CSMOKING_CrudePrev', 'BPHIGH_CrudePrev', 'OBESITY_CrudePrev', 'HIGHCHOL_CrudePrev')
established_demographics <- c('male_over_65_pct', 'female_over_65_pct')

established_disease_se <- c('CANCER_SE', 'ARTHRITIS_SE',  'STROKE_SE', 'CASTHMA_SE','COPD_SE', 'CHD_SE', 'DIABETES_SE', 'KIDNEY_SE', 'BPMED_SE')
established_risk_se <- c( 'CSMOKING_SE', 'BPHIGH_SE', 'OBESITY_SE', 'HIGHCHOL_SE')

filter_crudeprev <- function(arr) {arr[grep('CrudePrev',arr)]}
```

# Built environment predictors
```{r DL pipeline}
built_env <- read_csv('./health_inidicators_data_for_COVID_Tracts_XY_Model.csv')
built_env <- built_env %>% select(-X1)
built_env_city <- built_env %>% select(category, city, state, r2, rmse) %>% unique()
built_env_city_avg <- built_env %>% select(category, city, state, r2, rmse) %>% group_by(city, state) %>% summarize(avg_r2=mean(r2, na.rm = T), avg_rmse=mean(rmse, na.rm = T))
```

# Correlation between health indicators
## "Established" co-morbids for COVID-19
```{r correlation of index}
established_cols <- c(established_disease, established_risk, established_demographics)
non_na <- complete.cases(fh_acs[, established_cols])
fh_acs_established <- fh_acs[non_na, established_cols]

cr_established <- cor(fh_acs_established, use='pairwise.complete.obs')
heatmap.2(cr_established)
```




# PC of established COVID risk factors
```{r prcomp}
pc_established <- prcomp(fh_acs_established,center=T, scale.=T)
summary(pc_established)
pc_established$rotation
```


# PC1 and 2 explain 90% of variation
```{r prcomp 2}
plot(pc_established$x[, 1], fh_acs_established[, 'BPHIGH_CrudePrev'])
plot(pc_established$x[, 1], fh_acs_established[, 'BPMED_CrudePrev'])
plot(pc_established$x[, 1], fh_acs_established[, 'OBESITY_CrudePrev'])
plot(pc_established$x[, 1], fh_acs_established[, 'ARTHRITIS_CrudePrev'])
plot(pc_established$x[, 2], fh_acs_established[, 'CANCER_CrudePrev'])
fh_acs_established$pc_1 <- as.numeric(pc_established$x[, 1])
fh_acs_established$pc_2 <- as.numeric(pc_established$x[, 2])
```

```{r covid risk scores}
# now build a database of scores
# PC1 of all
# single
# age
placenames_cols <- c("stateabbr","placename" , "county_code", "state_code", "geoid","fips_place_tract","lat","long")
fh_acs_scoring <- fh_acs_established %>% cbind(fh_acs[non_na, placenames_cols])

fh_acs_scoring <- fh_acs_scoring %>% mutate(avg_over_65_pct = (male_over_65_pct+female_over_65_pct)/2)
fh_acs_scoring <- fh_acs_scoring %>% mutate(risk_score_pc_1=rescale(pc_1, to=c(0,100)))
fh_acs_scoring <- fh_acs_scoring %>% mutate(risk_score_pc_2=rescale(-pc_2, to=c(0,100))) # flip sign of pc2
fh_acs_scoring <- fh_acs_scoring %>% mutate(risk_score_both_pc=rescale(risk_score_pc_1*.61 + risk_score_pc_2*.24, to=c(0,100))) # rescale according the the contribution of the PC
fh_acs_scoring <- fh_acs_scoring %>% mutate(risk_score_all=rowSums(.[established_cols])) %>% mutate(risk_score_all=rescale(risk_score_all, to=c(0,100)))
```

# Examine PCs and association with an additive risk score
```{r examine PC}
top_per_state_pc1 <- fh_acs_scoring %>% group_by(stateabbr) %>% top_n(1,risk_score_pc_1) %>% ungroup() %>% filter(risk_score_pc_1 >= 50)
bottom_per_state_pc1 <- fh_acs_scoring %>% group_by(stateabbr) %>% top_n(1,-risk_score_pc_1) %>% ungroup() %>% filter(risk_score_pc_1 <= 20)

p <- ggplot(fh_acs_scoring, aes(risk_score_pc_1, risk_score_pc_2)) 
p <- p + geom_text_repel(data=top_per_state_pc1, aes(risk_score_pc_1, risk_score_pc_2, label=paste(placename, stateabbr)), size=3)
p <- p + geom_text_repel(data=bottom_per_state_pc1, aes(risk_score_pc_1, risk_score_pc_2, label=paste(placename, stateabbr)), size=3)
p <- p + geom_point(alpha=0.5, color='gray')
p

p <- ggplot(fh_acs_scoring, aes(risk_score_pc_1, risk_score_all)) 
p <- p + geom_point(alpha=0.5, color='gray')
p


p <- ggplot(fh_acs_scoring, aes(risk_score_both_pc, risk_score_pc_1)) 
p <- p + geom_point(alpha=0.5, color='gray')
p

p <- ggplot(fh_acs_scoring, aes(risk_score_both_pc, risk_score_pc_2)) 
p <- p + geom_point(alpha=0.5, color='gray')
p


p <- ggplot(fh_acs_scoring, aes(risk_score_both_pc, CANCER_CrudePrev)) 
p <- p + geom_point(alpha=0.5, color='gray')
p

p <- ggplot(fh_acs_scoring, aes(risk_score_both_pc, DIABETES_CrudePrev)) 
p <- p + geom_point(alpha=0.5, color='gray')
p

p <- ggplot(fh_acs_scoring, aes(risk_score_both_pc, CHD_CrudePrev)) 
p <- p + geom_point(alpha=0.5, color='gray')
p


summary(lm(risk_score_all ~ risk_score_pc_1, data=fh_acs_scoring))
summary(lm(risk_score_all ~ risk_score_pc_1, data=fh_acs_scoring))
summary(lm(CANCER_CrudePrev ~ risk_score_both_pc, data=fh_acs_scoring))

p <- ggplot(fh_acs_scoring, aes(risk_score_both_pc, avg_over_65_pct)) 
p <- p + geom_point(alpha=0.5, color='gray')
p

```



```{r}

top_per_state <- fh_acs_scoring %>% group_by(stateabbr) %>% top_n(1,risk_score_both_pc) %>% ungroup() %>% filter(risk_score_both_pc >= 60)
# top census tracts per state based on age, but filtered for those that have a high risk score
oldest_tracts <- fh_acs_scoring %>% group_by(stateabbr) %>% top_n(1,avg_over_65_pct) %>% ungroup() %>% filter(avg_over_65_pct >= .50)
top_both <- fh_acs_scoring %>% filter(risk_score_both_pc >= 60 & avg_over_65_pct >= .50)
p <- ggplot(fh_acs_scoring, aes(avg_over_65_pct, risk_score_both_pc))
p <- p + geom_point(alpha=0.5, color='gray')
p <- p + geom_point(data=top_per_state, aes(avg_over_65_pct, risk_score_both_pc))
p <- p + geom_text_repel(data=top_per_state, aes(avg_over_65_pct, risk_score_both_pc, label=paste(placename, stateabbr)), size=3)
p <- p + geom_point(data=oldest_tracts, aes(avg_over_65_pct, risk_score_both_pc), color='red')
p <- p + geom_text_repel(data=oldest_tracts, aes(avg_over_65_pct, risk_score_both_pc, label=paste(placename, stateabbr)),color='red', size=3)
p <- p + theme_fivethirtyeight() + theme(axis.title = element_text(), legend.position = 'none') + labs(x = '% Age over 65', y = 'Comorbidity Risk Score [PCs]')
p


fh_acs_scoring %>% arrange(desc(risk_score_both_pc)) %>% select(risk_score_both_pc, risk_score_pc_1, DIABETES_CrudePrev, CHD_CrudePrev, CANCER_CrudePrev, avg_over_65_pct,placename, stateabbr) %>% head(n=10)

```



# Randomization test
- see error_simulation.R

# Correlate with SES/deprivation and built environment
```{r}
ses_vars_to_bind <- setdiff(ses_vars, names(fh_acs_scoring))
fh_all <- cbind(fh_acs_scoring, fh_acs[non_na, ses_vars_to_bind])

fh_all_built <- fh_all %>% left_join(built_env_city_avg, by=c('placename'='city'))

p_age <- ggplot(fh_all, aes(I((male_over_65_pct + female_over_65_pct)/2), risk_score_both_pc))
p_age <- p_age + geom_point(alpha=0.5, color='gray') + xlab('% over 65') + ylab('Covid Community Risk Score')

p_income <- ggplot(fh_all, aes(median_income, risk_score_both_pc))
p_income <- p_income + geom_point(alpha=0.5, color='gray') + xlab('Median income') + ylab('Covid Community Risk Score')

p_poverty <- ggplot(fh_all, aes(at_below_poverty_pct, risk_score_both_pc))
p_poverty <- p_poverty + geom_point(alpha=0.5, color='gray') + xlab('% below poverty') + ylab('Covid Community Risk Score')

p_num_people <- ggplot(fh_all, aes(more_than_one_occupant_per_room_pct, risk_score_both_pc))
p_num_people <- p_num_people + geom_point(alpha=0.5, color='gray') + xlab('% of homes with >1 occupant per room') + ylab('Covid Community Risk Score')

#p_built_e <- ggplot(fh_all_built, aes(avg_r2, risk_score_pc_1))
#p_built_e <- p_built_e + geom_point(alpha=0.5) + xlab('Average(R2) of Built Environment Predictions') + ylab('Covid Community Risk Score')

ggarrange(p_age, p_income, p_poverty, p_num_people, ncol=2, nrow=2)

```


# Census-tract Community Risk data dump
```{r}
fh_acs_scoring <- fh_acs_scoring %>% cbind(fh_acs[non_na, c('total_males', 'total_females')])
fh_acs_scoring <- fh_acs_scoring %>% mutate(risk_score = risk_score_both_pc)
fh_acs_scoring <- fh_acs_scoring %>% group_split(placename) %>% map(function(.x) { .x %>% mutate(rank_in_city=rank(-risk_score)) }) %>% bind_rows()
fh_acs_scoring <- fh_acs_scoring %>% group_split(stateabbr) %>% map(function(.x) { .x %>% mutate(rank_in_state=rank(-risk_score)) }) %>% bind_rows()
write_rds(fh_acs_scoring, path = "fh_acs_covid_comm_score.rds")
write_csv(fh_acs_scoring, path="fh_acs_covid_comm_score.csv")
```

# City COVID-19 Community Risk Score

```{r city risk score}
fh_acs_scoring_city <- fh_acs_scoring %>% select(-c(pc_1, pc_2, risk_score_pc_1, risk_score_pc_2, risk_score, risk_score_both_pc, risk_score_all, lat,long, geoid, fips_place_tract)) %>% unite(place_state, placename, stateabbr, sep="|")
fh_acs_scoring_city <- fh_acs_scoring_city %>% mutate(total_population = total_males + total_females)


cols_to_avg <- c(established_disease, established_risk)
cols_se <- c(established_disease_se, established_risk_se)

fh_acs_scoring_city <- fh_acs_scoring_city %>% cbind(fh_acs[non_na, c(established_disease_se, established_risk_se)])

cols_to_total <- c("male_over_65_pct","female_over_65_pct","avg_over_65_pct")
cols_total <- c("male_over_65_count","female_over_65_count","avg_over_65_count")

for(i in 1:length(cols_to_total)) {
  fh_acs_scoring_city <- fh_acs_scoring_city %>% mutate(!!cols_total[i] := .data[[cols_to_total[i]]]*total_population)
}
fh_acs_scoring_by_city <- fh_acs_scoring_city %>% group_by(place_state) %>% summarise_at(c(cols_total, 'total_population'), sum)

cols_to_prev <- c("male_over_65_pct","female_over_65_pct","avg_over_65_pct")
for(i in 1:length(cols_total)) {
  fh_acs_scoring_by_city <- fh_acs_scoring_by_city %>% mutate(!!cols_to_prev[i] := .data[[cols_total[i]]]/total_population)
}


## now take weighted prevalence average
weighted_prev_avg <- function(d, prefix_col) { 
  # d is grouped by frame for a city or state
  prev_col <- sprintf('%s_CrudePrev', prefix_col)
  se_col <- sprintf('%s_SE', prefix_col)
  ind <- complete.cases(d[, c(prev_col, se_col)])
  d <- d[ind,]
  weighted.mean(d[[prev_col]], 1/(d[[se_col]]+0.0005)) # note the fudge factor
}


weighted_prev_avgs <- function(d, prefix_cols) {
  avgs <- prefix_cols %>% map_dbl(~weighted_prev_avg(d, .x))
  place<- d[['place_state']][1]
  tibble(disease=prefix_cols, prevalence=avgs, place_state=place)
}

prefixes <- c("CANCER", "ARTHRITIS", "STROKE","CASTHMA", "COPD", "CHD", "DIABETES","KIDNEY", "BPMED", "CSMOKING", "BPHIGH","OBESITY","HIGHCHOL" )

fh_acs_scoring_prevs_by_city <- fh_acs_scoring_city %>% group_by(place_state) %>% group_split(keep=TRUE) %>% map_df(~weighted_prev_avgs(.x, prefixes))
fh_acs_scoring_prevs_by_city$disease <- sprintf('%s_CrudePrev', fh_acs_scoring_prevs_by_city$disease)
fh_acs_scoring_prevs_by_city <- fh_acs_scoring_prevs_by_city %>% spread(disease, prevalence)

fh_acs_scoring_by_city <- fh_acs_scoring_by_city %>% left_join(fh_acs_scoring_prevs_by_city, by='place_state')


pc_by_city <- prcomp(fh_acs_scoring_by_city[,established_cols],center=T, scale.=T)
summary(pc_by_city)
pc_by_city$rotation

fh_acs_scoring_by_city$pc_1 <- as.numeric(pc_by_city$x[, 1])
fh_acs_scoring_by_city$pc_2 <- as.numeric(pc_by_city$x[, 2])

fh_acs_scoring_by_city <- fh_acs_scoring_by_city %>% mutate(risk_score_pc_1=as.numeric(rescale(pc_1, to=c(0,100)))) # flip the sign
fh_acs_scoring_by_city <- fh_acs_scoring_by_city %>% mutate(risk_score_pc_2=as.numeric(rescale(-pc_2, to=c(0,100))))
fh_acs_scoring_by_city <- fh_acs_scoring_by_city %>% mutate(risk_score_both_pc=rescale(risk_score_pc_1*.61 + risk_score_pc_2*.24, to=c(0,100))) # rescale according the the contribution of the PC
fh_acs_scoring_by_city <- fh_acs_scoring_by_city %>% mutate(risk_score_all=rowSums(.[established_cols])) %>% mutate(risk_score_all=as.numeric(rescale(risk_score_all, to=c(0,100))))

plot(fh_acs_scoring_by_city$risk_score_pc_1, fh_acs_scoring_by_city$risk_score_all)
plot(fh_acs_scoring_by_city$risk_score_pc_2, fh_acs_scoring_by_city$risk_score_all)
# add stateabbr
fh_acs_scoring_by_city <- fh_acs_scoring_by_city %>% separate(place_state, c("placename", "stateabbr"), sep="\\|")
  

top_per_state <- fh_acs_scoring_by_city %>% group_by(stateabbr) %>% top_n(5,risk_score_pc_1) %>% ungroup() 
DT::datatable(top_per_state %>% select(stateabbr, placename, risk_score_pc_1, risk_score_both_pc, risk_score_all))

fh_acs_scoring_by_city_distribute <- fh_acs_scoring_by_city %>% select(c(established_cols, "placename", "stateabbr", "risk_score_both_pc", "risk_score_pc_1", "risk_score_pc_2", "risk_score_all", "total_population")) 

fh_acs_scoring_by_city_distribute <- fh_acs_scoring_by_city_distribute %>% mutate(risk_score = risk_score_both_pc)
fh_acs_scoring_by_city_distribute <- fh_acs_scoring_by_city_distribute %>% group_split(stateabbr) %>% map(function(.x) { .x %>% mutate(rank_in_state=rank(-risk_score)) }) %>% bind_rows()

write_rds(fh_acs_scoring_by_city_distribute, path = "fh_acs_covid_city_comm_score.rds")
write_csv(fh_acs_scoring_by_city_distribute, path="fh_acs_covid_city_comm_score.csv")
```

# County COVID-19 Community Risk Score
```{r county risk score}
fh_acs_scoring_county <- fh_acs_scoring %>% select(-c(pc_1, pc_2, risk_score_pc_1, risk_score_pc_2, risk_score_all, risk_score_both_pc, lat,long, geoid, fips_place_tract)) %>% mutate(total_population = total_males + total_females)
location_info <- fh_acs_scoring %>% select(county_code, stateabbr) %>% unique() # go up a hierarchy
fh_acs_scoring_county <- fh_acs_scoring_county %>% cbind(fh_acs[non_na, c(established_disease_se, established_risk_se)])

for(i in 1:length(cols_to_total)) {
  fh_acs_scoring_county <- fh_acs_scoring_county %>% mutate(!!cols_total[i] := .data[[cols_to_total[i]]]*total_population)
}
fh_acs_scoring_by_county <- fh_acs_scoring_county %>% group_by(county_code) %>% summarise_at(c(cols_total, 'total_population'), sum)

for(i in 1:length(cols_total)) {
  fh_acs_scoring_by_county <- fh_acs_scoring_by_county %>% mutate(!!cols_to_prev[i] := .data[[cols_total[i]]]/total_population)
}

weighted_county_code_prev_avgs <- function(d, prefix_cols) {
  avgs <- prefix_cols %>% map_dbl(~weighted_prev_avg(d, .x))
  place<- d[['county_code']][1]
  tibble(disease=prefix_cols, prevalence=avgs, county_code=place)
}

fh_acs_scoring_prevs_by_county <- fh_acs_scoring_county %>% group_by(county_code) %>% group_split(keep=TRUE) %>% map_df(~weighted_county_code_prev_avgs(.x, prefixes))
fh_acs_scoring_prevs_by_county$disease <- sprintf('%s_CrudePrev', fh_acs_scoring_prevs_by_county$disease)
fh_acs_scoring_prevs_by_county <- fh_acs_scoring_prevs_by_county %>% spread(disease, prevalence)

fh_acs_scoring_by_county <- fh_acs_scoring_by_county %>% left_join(fh_acs_scoring_prevs_by_county, by='county_code')

heatmap.2(cor(fh_acs_scoring_by_county[,established_cols]))

pc_by_county <- prcomp(fh_acs_scoring_by_county[,established_cols],center=T, scale.=T)
summary(pc_by_county)
pc_by_county$rotation


fh_acs_scoring_by_county$pc_1 <- as.numeric(pc_by_county$x[, 1])
fh_acs_scoring_by_county$pc_2 <- as.numeric(pc_by_county$x[, 2])


fh_acs_scoring_by_county <- fh_acs_scoring_by_county %>% mutate(risk_score_pc_1=as.numeric(rescale(-pc_1,to=c(0,100)))) 
fh_acs_scoring_by_county <- fh_acs_scoring_by_county %>% mutate(risk_score_pc_2=as.numeric(rescale(-pc_2,to=c(0,100))))
fh_acs_scoring_by_county <- fh_acs_scoring_by_county %>% mutate(risk_score_both_pc=rescale(risk_score_pc_1*.61 + risk_score_pc_2*.24, to=c(0,100))) # rescale according the the contribution of the PC
fh_acs_scoring_by_county <- fh_acs_scoring_by_county %>% mutate(risk_score_all=rowSums(.[established_cols])) %>% mutate(risk_score_all=as.numeric(rescale(risk_score_all, to=c(0,100))))

plot(fh_acs_scoring_by_county$risk_score_pc_1, fh_acs_scoring_by_county$risk_score_all)
plot(fh_acs_scoring_by_county$risk_score_pc_2, fh_acs_scoring_by_county$risk_score_all)

fh_acs_scoring_by_county_distribute <- fh_acs_scoring_by_county %>% select(c(established_cols, "county_code", "risk_score_both_pc", "risk_score_pc_1", "risk_score_pc_2", "risk_score_all", "total_population")) 

fh_acs_scoring_by_county_distribute <- fh_acs_scoring_by_county_distribute %>% mutate(risk_score = risk_score_both_pc)
fh_acs_scoring_by_county_distribute <- fh_acs_scoring_by_county_distribute %>% left_join(location_info)
fh_acs_scoring_by_county_distribute <- fh_acs_scoring_by_county_distribute %>% group_split(stateabbr) %>% map(function(.x) { .x %>% mutate(rank_in_state=rank(-risk_score)) }) %>% bind_rows()

DT::datatable(fh_acs_scoring_by_county_distribute %>% select(county_code, stateabbr, risk_score, risk_score_pc_1, risk_score_all))
write_rds(fh_acs_scoring_by_county_distribute, path = "fh_acs_covid_county_comm_score.rds")
write_csv(fh_acs_scoring_by_county_distribute, path="fh_acs_covid_county_comm_score.csv")
```




# State COVID-19 Community Risk Score
```{r state risk score}
fh_acs_scoring_state <- fh_acs_scoring %>% select(-c(pc_1, pc_2, risk_score_pc_1, risk_score_pc_2, risk_score_all, risk_score_both_pc, lat,long, geoid, fips_place_tract)) %>% mutate(total_population = total_males + total_females)

fh_acs_scoring_state <- fh_acs_scoring_state %>% cbind(fh_acs[non_na, c(established_disease_se, established_risk_se)])

cols_to_total <- c("male_over_65_pct","female_over_65_pct","avg_over_65_pct")
cols_total <- c("male_over_65_count","female_over_65_count","avg_over_65_count")

for(i in 1:length(cols_to_total)) {
  fh_acs_scoring_state <- fh_acs_scoring_state %>% mutate(!!cols_total[i] := .data[[cols_to_total[i]]]*total_population)
}
fh_acs_scoring_by_state <- fh_acs_scoring_state %>% group_by(stateabbr) %>% summarise_at(c(cols_total, 'total_population'), sum)

cols_to_prev <- c("male_over_65_pct","female_over_65_pct","avg_over_65_pct")
for(i in 1:length(cols_total)) {
  fh_acs_scoring_by_state <- fh_acs_scoring_by_state %>% mutate(!!cols_to_prev[i] := .data[[cols_total[i]]]/total_population)
}

weighted_state_prev_avgs <- function(d, prefix_cols) {
  avgs <- prefix_cols %>% map_dbl(~weighted_prev_avg(d, .x))
  place<- d[['stateabbr']][1]
  tibble(disease=prefix_cols, prevalence=avgs, stateabbr=place)
}

fh_acs_scoring_prevs_by_state <- fh_acs_scoring_state %>% group_by(stateabbr) %>% group_split(keep=TRUE) %>% map_df(~weighted_state_prev_avgs(.x, prefixes))
fh_acs_scoring_prevs_by_state$disease <- sprintf('%s_CrudePrev', fh_acs_scoring_prevs_by_state$disease)
fh_acs_scoring_prevs_by_state <- fh_acs_scoring_prevs_by_state %>% spread(disease, prevalence)

fh_acs_scoring_by_state <- fh_acs_scoring_by_state %>% left_join(fh_acs_scoring_prevs_by_state, by='stateabbr')

heatmap.2(cor(fh_acs_scoring_by_state[,established_cols]))

pc_by_state <- prcomp(fh_acs_scoring_by_state[,established_cols],center=T, scale.=T)
summary(pc_by_state)
pc_by_state$rotation


fh_acs_scoring_by_state$pc_1 <- as.numeric(pc_by_state$x[, 1])
fh_acs_scoring_by_state$pc_2 <- as.numeric(pc_by_state$x[, 2])

fh_acs_scoring_by_state <- fh_acs_scoring_by_state %>% mutate(risk_score_pc_1=as.numeric(rescale(pc_1,to=c(0,100))))
fh_acs_scoring_by_state <- fh_acs_scoring_by_state %>% mutate(risk_score_pc_2=as.numeric(rescale(-pc_2,to=c(0,100))))
fh_acs_scoring_by_state <- fh_acs_scoring_by_state %>% mutate(risk_score_both_pc=rescale(risk_score_pc_1*.61 + risk_score_pc_2*.24, to=c(0,100))) # rescale according the the contribution of the PC
fh_acs_scoring_by_state <- fh_acs_scoring_by_state %>% mutate(risk_score_all=rowSums(.[established_cols])) %>% mutate(risk_score_all=as.numeric(rescale(risk_score_all,to=c(0,100))))

#plot(fh_acs_scoring_by_state$risk_score_pc_1, fh_acs_scoring_by_state$risk_score_all)
#plot(fh_acs_scoring_by_state$risk_score_pc_2, fh_acs_scoring_by_state$risk_score_all)

fh_acs_scoring_by_state_distribute <- fh_acs_scoring_by_state %>% select(c(established_cols, "stateabbr", "risk_score_both_pc", "risk_score_pc_1", "risk_score_pc_2", "risk_score_all", "total_population")) 

fh_acs_scoring_by_state_distribute <- fh_acs_scoring_by_state_distribute %>% mutate(rank_in_country=rank(risk_score_both_pc)) %>% mutate(risk_score = risk_score_both_pc)

DT::datatable(fh_acs_scoring_by_state_distribute %>% select(stateabbr, risk_score,risk_score_all))
write_rds(fh_acs_scoring_by_state_distribute, path = "fh_acs_covid_state_comm_score.rds")
write_csv(fh_acs_scoring_by_state_distribute, path="fh_acs_covid_state_comm_score.csv")

```


