---
output:
  html_document:
    df_print: paged
---

# Covid-19 Community Score
- finished: 04/02/2020
- minor update 09/20/20
- chirag@xy.ai


```{r library load}
library(ggpubr)
library(tidyverse)
library(readr)
library(gplots)
library(ggrepel)
library(ggthemes)
library(DT)
library(scales)
library(corrr)
library(randomForest)
WRITE_OUTPUT <- FALSE
RUN_RF <- FALSE
```

```{r load and prepare data}
source('data_prep_risk_score.R') ## 
```

# Correlation between health indicators
## "Established" co-morbids for COVID-19
```{r correlation of index}
cr_established <- cor(fh_acs_established, use='pairwise.complete.obs')

new_names <- c('Cancer', 'Arthritis', 'Stroke', 'Chronic Asthma', 'COPD', 'Heart Disease', 'Diabetes', 'Kidney Disease', 'BP Medication', 'Smoking', 'High BP', 'Obesity', 'High Cholesterol', 'Male Over 65', 'Female Over 65')

colnames(cr_established) <- new_names
rownames(cr_established) <- new_names
heatmapColors <- function(numColors=16) {
	c1 <- rainbow(numColors,v=seq(0.5,1,length=numColors),s=seq(1,0.3,length=numColors),start=4/6,end=4.0001/6);
	c2 <- rainbow(numColors,v=seq(0.5,1,length=numColors),s=seq(1,0.3,length=numColors),start=1/6,end=1.0001/6);
	c3 <- c(c1,rev(c2)); 
	return(c3)
}
heatmap.2(cr_established, trace='none',margins = c(10, 10), col=heatmapColors(8))

quantile(abs(cr_established[upper.tri(cr_established)]))

fh_acs_established_2 <- fh_acs_established
colnames(fh_acs_established_2) <- new_names
cr_established_tidy <- fh_acs_established_2 %>% correlate() %>% stretch()
diabetes_related <- c('Heart Disease', 'Diabetes', 'Kidney Disease', 'Stroke')
risk_factors_diabetes <- c('High BP', 'Obesity', 'High Cholesterol')

                  
cr_established_tidy %>% filter(x %in% diabetes_related, y %in% diabetes_related) %>% summarize(quantile_low=quantile(r, na.rm = T, probs=.25), median=median(r, na.rm=T), quantile_high=quantile(r, na.rm=T, probs=.75), avg=mean(abs(r), na.rm=T))

cr_established_tidy %>% filter(x %in% risk_factors_diabetes, y %in% risk_factors_diabetes) %>% summarize(quantile_low=quantile(r, na.rm = T, probs=.25), median=median(r, na.rm=T), quantile_high=quantile(r, na.rm=T, probs=.75), avg=mean(abs(r), na.rm=T))

cr_established_tidy %>% filter(x %in% risk_factors_diabetes, y %in% diabetes_related) %>% summarize(quantile_low=quantile(r, na.rm = T, probs=.25), median=median(r, na.rm=T), quantile_high=quantile(r, na.rm=T, probs=.75), avg=mean(abs(r), na.rm=T))

cr_established_tidy %>% filter(x == 'Chronic Asthma', y == 'COPD')
cr_established_tidy %>% filter(x == 'Smoking', y == 'COPD')
cr_established_tidy %>% filter(x == 'Smoking', y == 'Chronic Asthma')
cr_established_tidy %>% filter(x == 'Obesity') %>% summarize(mean=mean(abs(r), na.rm=T))
cr_established_tidy %>% filter(x == 'Cancer', y %in% c('Male Over 65', 'Female Over 65'))  %>% summarize(mean=mean(abs(r), na.rm=T))
remove(fh_acs_established_2)
```

```{r raw prevalence across census tracts, fig.height=8, fig.width=8}
# number of tracts
num_tracts <- fh_acs[non_na, c("placename", "county_code", "stateabbr", "total_population")] 
num_tracts %>% group_by(placename, stateabbr) %>% summarize(number_of_tracts=n(), total_population=sum(total_population)) %>% arrange(desc(number_of_tracts))

num_tracts %>% group_by(placename, stateabbr) %>% summarize(number_of_tracts=n()) %>% ungroup() %>% summarise(median=median(number_of_tracts), low=min(number_of_tracts), hi=max(number_of_tracts), pct_25=quantile(number_of_tracts, probs=.25), pct_75=quantile(number_of_tracts, probs = .75),sd=sd(number_of_tracts)) %>% mutate(IQR = pct_75 -pct_25)


fh_acs_established_long <- fh_acs_established %>% cbind(fh_acs[non_na,"placename"])
names(fh_acs_established_long) <- c(new_names, "placename")
fh_acs_established_long <-  fh_acs_established_long %>% gather('disease', 'prevalence', -placename)

fh_acs_established_long <- fh_acs_established_long %>% mutate(disease = fct_reorder(disease, prevalence, .fun='median'))
prevalence_p <- ggplot(fh_acs_established_long, aes(disease, prevalence*100)) 
prevalence_p <- prevalence_p + geom_violin()
prevalence_p <- prevalence_p + xlab('') + ylab('Prevalence') + coord_flip() + theme_fivethirtyeight() + theme(axis.title = element_text())
prevalence_p

DT::datatable(fh_acs_established_long %>% group_by(disease) %>% summarize(median=median(prevalence), low=min(prevalence), hi=max(prevalence), pct_25=quantile(prevalence, probs=.25), pct_75=quantile(prevalence, probs = .75), sd=sd(prevalence)) %>% arrange(desc(median)))

DT::datatable(fh_acs_established_long %>% group_by(disease, placename) %>% summarise(median=median(prevalence), low=min(prevalence), hi=max(prevalence), pct_25=quantile(prevalence, probs=.25), pct_75=quantile(prevalence, probs = .75), sd=sd(prevalence)) %>% arrange(desc(sd)))


summ_prev_per_city <- fh_acs_established_long %>% group_by(disease, placename) %>% summarise(median=median(prevalence), low=min(prevalence), hi=max(prevalence), pct_25=quantile(prevalence, probs=.25), pct_75=quantile(prevalence, probs = .75),sd=sd(prevalence)) %>% mutate(IQR = pct_75 -pct_25)

## plot the IQR vs. median prevalence per city

summ_prev_top_iqrs <- summ_prev_per_city %>% group_by(disease) %>% top_n(3, IQR) %>% ungroup()
p <- ggplot(summ_prev_per_city, aes(median, IQR)) 
p <- p + geom_point() + facet_wrap(~disease, nrow=3)
p <- p + geom_text_repel(data=summ_prev_top_iqrs, aes(median, IQR, label=placename), color='red', size=3)
p <- p + geom_point(data=summ_prev_top_iqrs, aes(median, IQR), color='red')
p <- p + theme_fivethirtyeight() + theme(axis.title = element_text(), legend.position = 'none') + labs(x='Median Prevalence in a City', y='75th - 25th Percentile of Prevalence in a City')
p

```




# PC of established COVID risk factors
```{r prcomp}
pc_established <- prcomp(fh_acs_established,center=T, scale.=T)
summary(pc_established)
pc_established$rotation [, c(1,2)]
```


# PC1 and 2 explain 90% of variation
```{r prcomp 2}
fh_acs_established$pc_1 <- as.numeric(pc_established$x[, 1])
fh_acs_established$pc_2 <- as.numeric(pc_established$x[, 2])
```

```{r covid risk scores}
placenames_cols <- c("stateabbr","placename" , "county_code", "state_code", "geoid","fips_place_tract","lat","long")
fh_acs_scoring <- fh_acs_established %>% cbind(fh_acs[non_na, placenames_cols])

fh_acs_scoring <- fh_acs_scoring %>% mutate(avg_over_65_pct = (male_over_65_pct+female_over_65_pct)/2)
fh_acs_scoring <- fh_acs_scoring %>% mutate(risk_score_pc_1=rescale(pc_1, to=c(0,100)))
fh_acs_scoring <- fh_acs_scoring %>% mutate(risk_score_pc_2=rescale(-pc_2, to=c(0,100))) # flip sign of pc2
fh_acs_scoring <- fh_acs_scoring %>% mutate(risk_score_both_pc=rescale(risk_score_pc_1*.61 + risk_score_pc_2*.24, to=c(0,100))) # rescale according the the contribution of the PC
fh_acs_scoring <- fh_acs_scoring %>% mutate(risk_score_all=rowSums(.[established_cols])) %>% mutate(risk_score_all=rescale(risk_score_all, to=c(0,100)))

## load in the errors 
comm_score_error <- read_rds('./scores/fh_acs_covid_comm_score_error.rds') # see error_simulation.R
fh_acs_scoring <- fh_acs_scoring %>% mutate(score_error=comm_score_error$score_error)
fh_acs_scoring %>% arrange(score_error) %>% dplyr::select(c(score_error, placename)) 

```

# Examine PCs and association with an additive risk score
```{r  examine PC, fig.width=8, fig.height=8}

# show the top 5 tracts for the each axis
fh_acs_scoring_top <- fh_acs_scoring %>% top_n(10, risk_score_pc_1 ) %>%
  rbind(fh_acs_scoring %>% top_n(10, risk_score_pc_2 )) %>% unite(place, placename, stateabbr, sep=",")

p <- ggplot(fh_acs_scoring, aes(risk_score_pc_1, risk_score_pc_2)) 
p <- p + geom_point(alpha=0.5, color='gray')
p <- p + geom_point(data=fh_acs_scoring_top, aes(risk_score_pc_1, risk_score_pc_2), color='red')
p <- p + geom_text_repel(data=fh_acs_scoring_top, aes(risk_score_pc_1, risk_score_pc_2, label=place), color='red',size=3)
p <- p + theme_fivethirtyeight() + theme(axis.title = element_text(), legend.position = 'none') + labs(x='Component 1 (61%)', y='Component 2 (24%)')
p



## now get the variation per city for the risk score
summarized_risk_score_by_city <- fh_acs_scoring %>% group_by(placename) %>% summarise(stateabbr=first(stateabbr), median=median(risk_score_both_pc), low=min(risk_score_both_pc), hi=max(risk_score_both_pc), pct_25=quantile(risk_score_both_pc, probs=.25), pct_75=quantile(risk_score_both_pc, probs = .75),sd=sd(risk_score_both_pc)) %>% mutate(IQR = pct_75 -pct_25)

summarized_risk_score_by_city_top_iqr <- summarized_risk_score_by_city %>% top_n(10, IQR)
summarized_risk_score_by_city_top_iqr <- summarized_risk_score_by_city_top_iqr %>% rbind(summarized_risk_score_by_city %>% filter(median >= 40, IQR > 10),summarized_risk_score_by_city %>% filter(median >= 50)) %>% unite(place, placename, stateabbr, sep=",")

p <- ggplot(summarized_risk_score_by_city, aes(median, IQR)) 
p <- p+ geom_point(alpha=.5)
p <- p + geom_text_repel(data=summarized_risk_score_by_city_top_iqr, aes(median, IQR, label=place), color='red', size=3)
p <- p + geom_point(data=summarized_risk_score_by_city_top_iqr, aes(median, IQR), color='red')
p <- p +  theme_fivethirtyeight() + theme(axis.title = element_text(), legend.position = 'none') + labs(x = '% Median C19 Risk Score in a City', y = 'Difference in IQR of C19 Risk Score in a City') 
p
summarized_risk_score_by_city_top_iqr 

```


# Correlation with Sociodemographic Factors

```{r ses}
ses_vars_to_bind <- setdiff(ses_vars, names(fh_acs_scoring))
fh_all <- cbind(fh_acs_scoring, fh_acs[non_na, ses_vars_to_bind])

top_per_state <- fh_acs_scoring %>% group_by(stateabbr) %>% top_n(1,risk_score_both_pc) %>% ungroup() %>% filter(risk_score_both_pc >= 75)
# top census tracts per state based on age, but filtered for those that have a high risk score
oldest_tracts <- fh_acs_scoring %>% group_by(stateabbr) %>% top_n(1,avg_over_65_pct) %>% ungroup() %>% filter(avg_over_65_pct >= .50)
top_both <- fh_acs_scoring %>% filter(risk_score_both_pc >= 75 & avg_over_65_pct >= .50)
p <- ggplot(fh_acs_scoring, aes(I(100*avg_over_65_pct), risk_score_both_pc))
p <- p + geom_point(alpha=0.5, color='gray')
p <- p + geom_point(data=top_per_state, aes(I(100*avg_over_65_pct), risk_score_both_pc))
p <- p + geom_text_repel(data=top_per_state, aes(I(100*avg_over_65_pct), risk_score_both_pc, label=paste(placename, stateabbr)), size=3)
p <- p + geom_point(data=oldest_tracts, aes(I(100*avg_over_65_pct), risk_score_both_pc), color='red')
p <- p + geom_text_repel(data=oldest_tracts, aes(I(100*avg_over_65_pct), risk_score_both_pc, label=paste(placename, stateabbr)),color='red', size=3)
p <- p + theme_fivethirtyeight() + theme(axis.title = element_text(), legend.position = 'none') + labs(x = '% Age over 65', y = '')
p


top_score <- fh_all %>% group_by(stateabbr) %>% top_n(2, risk_score_both_pc) %>% filter(risk_score_both_pc >= 75)
p_income <- ggplot(fh_all, aes(median_income, risk_score_both_pc))
p_income <- p_income + geom_point(alpha=0.5, color='gray') 
p_income <- p_income + geom_point(data=top_score, aes(median_income, risk_score_both_pc))
p_income <- p_income + geom_text_repel(data=top_score, aes(median_income, risk_score_both_pc, label=paste(placename, stateabbr)),color='red', size=3)
p_income <- p_income + theme_fivethirtyeight() + theme(axis.title = element_text(), legend.position = 'none') + labs(x = 'Median Income per Tract', y = '')
p_income


p_poverty <- ggplot(fh_all, aes(I(100*at_below_poverty_pct), risk_score_both_pc))
p_poverty <- p_poverty + geom_point(alpha=0.5, color='gray')
p_poverty <- p_poverty + geom_point(data=top_score, aes(I(100*at_below_poverty_pct), risk_score_both_pc))
p_poverty <- p_poverty + geom_text_repel(data=top_score, aes(I(100*at_below_poverty_pct), risk_score_both_pc, label=paste(placename, stateabbr)),color='red', size=3)

high_poverty <- fh_all %>% group_by(stateabbr) %>% top_n(1, at_below_poverty_pct) %>% filter(at_below_poverty_pct >= .60, risk_score_both_pc > 50)
p_poverty <- p_poverty + geom_point(data=high_poverty, aes(I(100*at_below_poverty_pct), risk_score_both_pc))
p_poverty <- p_poverty + geom_text_repel(data=high_poverty, aes(I(100*at_below_poverty_pct), risk_score_both_pc, label=paste(placename, stateabbr)),color='red', size=3)
p_poverty <- p_poverty + theme_fivethirtyeight() + theme(axis.title = element_text(), legend.position = 'none') + labs(x = '% Below Poverty per Tract', y = '')
p_poverty
```


```{r SES predictors of C19 Risk Score}
set.seed(123)
fh_all_train_ind <- sample(nrow(fh_all), size=floor(nrow(fh_all)/2))
train <- fh_all[fh_all_train_ind,]
test <- fh_all[setdiff(1:nrow(fh_all), fh_all_train_ind), ]

mod_linear <- lm(risk_score_both_pc ~ scale(median_income) + scale(median_home_value) + scale(at_below_poverty_pct) + scale(unemployment_pct) + scale(non_employment_pct) + scale(less_than_high_school_pct) + scale(no_health_insurance_pct) + scale(more_than_one_occupant_per_room_pct) + scale(african_american_pct) + scale(hispanic_pct) + scale(asian_pct) + scale(other_race_pct), data=train)

test_prediction_lm <- tibble(predicted=predict(mod_linear, test), actual=test$risk_score_both_pc)
predicted_cases_index <- complete.cases(test_prediction_lm)
test_prediction_lm <- test_prediction_lm[predicted_cases_index, ]
summary(lm(predicted ~ actual, test_prediction_lm))

if(RUN_RF) {
  
mod_rf <- randomForest(risk_score_both_pc ~ median_income + unemployment_pct + median_home_value + at_below_poverty_pct + unemployment_pct + non_employment_pct + less_than_high_school_pct + no_health_insurance_pct + more_than_one_occupant_per_room_pct + african_american_pct + hispanic_pct + asian_pct + other_race_pct, data=train, na.action=na.roughfix, ntree=1000, importance=T)
  importance(mod_rf)
  test_prediction_rf <- tibble(predicted=predict(mod_rf, test), actual=test$risk_score_both_pc)
predicted_cases_index <- complete.cases(test_prediction_rf)
test_prediction_rf <- test_prediction_rf[predicted_cases_index, ]
summary(lm(predicted ~ actual, test_prediction_rf))

p <- ggplot(test_prediction_rf, aes(actual, predicted))
p <- p + geom_point(alpha=.1)
p

pred <- predict(mod_rf, fh_all)
fh_all <- fh_all %>% mutate(risk_score_both_pc_residual = risk_score_both_pc-pred)
  if(WRITE_OUTPUT) {
    fh_all_ses_residual <- fh_all %>% dplyr::select(fips_place_tract, geoid, placename, stateabbr, risk_score_both_pc, risk_score_both_pc_residual)

    write_rds(fh_all_ses_residual, path = "fh_acs_ses_residual_covid_comm_score.rds")
   write_csv(fh_all_ses_residual, path="fh_acs_ses_residual_covid_comm_score.csv")
  }
}





```





# Census-tract Community Risk data dump
```{r}
fh_acs_scoring <- fh_acs_scoring %>% cbind(fh_acs[non_na, c('total_males', 'total_females')])
fh_acs_scoring <- fh_acs_scoring %>% mutate(risk_score = risk_score_both_pc)
fh_acs_scoring <- fh_acs_scoring %>% group_split(placename) %>% map(function(.x) { .x %>% mutate(rank_in_city=rank(-risk_score)) }) %>% bind_rows()
fh_acs_scoring <- fh_acs_scoring %>% group_split(stateabbr) %>% map(function(.x) { .x %>% mutate(rank_in_state=rank(-risk_score)) }) %>% bind_rows()
if(WRITE_OUTPUT) {
  write_rds(fh_acs_scoring, path = "fh_acs_covid_comm_score.rds")
  write_csv(fh_acs_scoring, path="fh_acs_covid_comm_score.csv")
}
```

# City COVID-19 Community Risk Score
```{r city risk score}
fh_acs_scoring_city <- fh_acs_scoring %>% dplyr::select(-c(pc_1, pc_2, risk_score_pc_1, risk_score_pc_2, risk_score, risk_score_both_pc, risk_score_all, lat,long, geoid, fips_place_tract)) %>% unite(place_state, placename, stateabbr, sep="|")
fh_acs_scoring_city <- fh_acs_scoring_city %>% mutate(total_population = total_males + total_females)


cols_to_avg <- c(established_disease, established_risk)
cols_se <- c(established_disease_se, established_risk_se)

fh_acs_scoring_city <- fh_acs_scoring_city %>% cbind(fh_acs[non_na, c(established_disease_se, established_risk_se)])

cols_to_total <- c("male_over_65_pct","female_over_65_pct","avg_over_65_pct")
cols_total <- c("male_over_65_count","female_over_65_count","avg_over_65_count")

for(i in 1:length(cols_to_total)) {
  fh_acs_scoring_city <- fh_acs_scoring_city %>% mutate(!!cols_total[i] := .data[[cols_to_total[i]]]*total_population)
}
fh_acs_scoring_by_city <- fh_acs_scoring_city %>% group_by(place_state) %>% summarise_at(c(cols_total, 'total_population'), sum)

cols_to_prev <- c("male_over_65_pct","female_over_65_pct","avg_over_65_pct")
for(i in 1:length(cols_total)) {
  fh_acs_scoring_by_city <- fh_acs_scoring_by_city %>% mutate(!!cols_to_prev[i] := .data[[cols_total[i]]]/total_population)
}


## now take weighted prevalence average
weighted_prev_avg <- function(d, prefix_col) { 
  # d is grouped by frame for a city or state
  prev_col <- sprintf('%s_CrudePrev', prefix_col)
  se_col <- sprintf('%s_SE', prefix_col)
  ind <- complete.cases(d[, c(prev_col, se_col)])
  d <- d[ind,]
  weighted.mean(d[[prev_col]], 1/(d[[se_col]]+0.0005)) # note the fudge factor
}


weighted_prev_avgs <- function(d, prefix_cols) {
  avgs <- prefix_cols %>% map_dbl(~weighted_prev_avg(d, .x))
  place<- d[['place_state']][1]
  tibble(disease=prefix_cols, prevalence=avgs, place_state=place)
}

prefixes <- c("CANCER", "ARTHRITIS", "STROKE","CASTHMA", "COPD", "CHD", "DIABETES","KIDNEY", "BPMED", "CSMOKING", "BPHIGH","OBESITY","HIGHCHOL" )

fh_acs_scoring_prevs_by_city <- fh_acs_scoring_city %>% group_by(place_state) %>% group_split(keep=TRUE) %>% map_df(~weighted_prev_avgs(.x, prefixes))
fh_acs_scoring_prevs_by_city$disease <- sprintf('%s_CrudePrev', fh_acs_scoring_prevs_by_city$disease)
fh_acs_scoring_prevs_by_city <- fh_acs_scoring_prevs_by_city %>% spread(disease, prevalence)

fh_acs_scoring_by_city <- fh_acs_scoring_by_city %>% left_join(fh_acs_scoring_prevs_by_city, by='place_state')


pc_by_city <- prcomp(fh_acs_scoring_by_city[,established_cols],center=T, scale.=T)
summary(pc_by_city)
pc_by_city$rotation

fh_acs_scoring_by_city$pc_1 <- as.numeric(pc_by_city$x[, 1])
fh_acs_scoring_by_city$pc_2 <- as.numeric(pc_by_city$x[, 2])

fh_acs_scoring_by_city <- fh_acs_scoring_by_city %>% mutate(risk_score_pc_1=as.numeric(rescale(-pc_1, to=c(0,100)))) # flip the sign
fh_acs_scoring_by_city <- fh_acs_scoring_by_city %>% mutate(risk_score_pc_2=as.numeric(rescale(-pc_2, to=c(0,100))))
fh_acs_scoring_by_city <- fh_acs_scoring_by_city %>% mutate(risk_score_both_pc=rescale(risk_score_pc_1*.61 + risk_score_pc_2*.24, to=c(0,100))) # rescale according the the contribution of the PC
fh_acs_scoring_by_city <- fh_acs_scoring_by_city %>% mutate(risk_score_all=rowSums(.[established_cols])) %>% mutate(risk_score_all=as.numeric(rescale(risk_score_all, to=c(0,100))))

plot(fh_acs_scoring_by_city$risk_score_pc_1, fh_acs_scoring_by_city$risk_score_all)
plot(fh_acs_scoring_by_city$risk_score_pc_2, fh_acs_scoring_by_city$risk_score_all)
# add stateabbr
plot(fh_acs_scoring_by_city$risk_score_both_pc, fh_acs_scoring_by_city$avg_over_65_pct)
plot(fh_acs_scoring_by_city$risk_score_both_pc, fh_acs_scoring_by_city$DIABETES_CrudePrev)
fh_acs_scoring_by_city <- fh_acs_scoring_by_city %>% separate(place_state, c("placename", "stateabbr"), sep="\\|")
  

top_per_state <- fh_acs_scoring_by_city %>% group_by(stateabbr) %>% top_n(5,risk_score_pc_1) %>% ungroup() 
DT::datatable(top_per_state %>% dplyr::select(stateabbr, placename, risk_score_pc_1, risk_score_both_pc, risk_score_all))

fh_acs_scoring_by_city_distribute <- fh_acs_scoring_by_city %>% dplyr::select(c(established_cols, "placename", "stateabbr", "risk_score_both_pc", "risk_score_pc_1", "risk_score_pc_2", "risk_score_all", "total_population")) 

fh_acs_scoring_by_city_distribute <- fh_acs_scoring_by_city_distribute %>% mutate(risk_score = risk_score_both_pc)
fh_acs_scoring_by_city_distribute <- fh_acs_scoring_by_city_distribute %>% group_split(stateabbr) %>% map(function(.x) { .x %>% mutate(rank_in_state=rank(-risk_score)) }) %>% bind_rows()
if(WRITE_OUTPUT) {
  write_rds(fh_acs_scoring_by_city, path = "fh_acs_covid_city_comm_score.rds")
  write_csv(fh_acs_scoring_by_city_distribute, path="fh_acs_covid_city_comm_score.csv")
}
```

# County COVID-19 Community Risk Score
```{r county risk score}
fh_acs_scoring_county <- fh_acs_scoring %>% dplyr::select(-c(pc_1, pc_2, risk_score_pc_1, risk_score_pc_2, risk_score_all, risk_score_both_pc, lat,long, geoid, fips_place_tract)) %>% mutate(total_population = total_males + total_females)
location_info <- fh_acs_scoring %>% dplyr::select(county_code, stateabbr) %>% unique() # go up a hierarchy
fh_acs_scoring_county <- fh_acs_scoring_county %>% cbind(fh_acs[non_na, c(established_disease_se, established_risk_se)])

county_counts <- fh_acs_scoring_county %>% group_by(county_code) %>% summarize(n=n())

fh_acs_scoring_county <- fh_acs_scoring_county %>% left_join(county_counts) %>% filter(n > 1)

for(i in 1:length(cols_to_total)) {
  fh_acs_scoring_county <- fh_acs_scoring_county %>% mutate(!!cols_total[i] := .data[[cols_to_total[i]]]*total_population)
}
fh_acs_scoring_by_county <- fh_acs_scoring_county %>% group_by(county_code) %>% summarise_at(c(cols_total, 'total_population'), sum)

for(i in 1:length(cols_total)) {
  fh_acs_scoring_by_county <- fh_acs_scoring_by_county %>% mutate(!!cols_to_prev[i] := .data[[cols_total[i]]]/total_population)
}

weighted_county_code_prev_avgs <- function(d, prefix_cols) {
  avgs <- prefix_cols %>% map_dbl(~weighted_prev_avg(d, .x))
  place<- d[['county_code']][1]
  tibble(disease=prefix_cols, prevalence=avgs, county_code=place)
}

fh_acs_scoring_prevs_by_county <- fh_acs_scoring_county %>% group_by(county_code) %>% group_split(keep=TRUE) %>% map_df(~weighted_county_code_prev_avgs(.x, prefixes))
fh_acs_scoring_prevs_by_county$disease <- sprintf('%s_CrudePrev', fh_acs_scoring_prevs_by_county$disease)
fh_acs_scoring_prevs_by_county <- fh_acs_scoring_prevs_by_county %>% spread(disease, prevalence)

fh_acs_scoring_by_county <- fh_acs_scoring_by_county %>% left_join(fh_acs_scoring_prevs_by_county, by='county_code')

heatmap.2(cor(fh_acs_scoring_by_county[,established_cols]))

pc_by_county <- prcomp(fh_acs_scoring_by_county[,established_cols],center=T, scale.=T)
summary(pc_by_county)
pc_by_county$rotation


fh_acs_scoring_by_county$pc_1 <- as.numeric(pc_by_county$x[, 1])
fh_acs_scoring_by_county$pc_2 <- as.numeric(pc_by_county$x[, 2])


fh_acs_scoring_by_county <- fh_acs_scoring_by_county %>% mutate(risk_score_pc_1=as.numeric(rescale(-pc_1,to=c(0,100)))) 
fh_acs_scoring_by_county <- fh_acs_scoring_by_county %>% mutate(risk_score_pc_2=as.numeric(rescale(-pc_2,to=c(0,100))))
fh_acs_scoring_by_county <- fh_acs_scoring_by_county %>% mutate(risk_score_both_pc=rescale(risk_score_pc_1*.61 + risk_score_pc_2*.24, to=c(0,100))) # rescale according the the contribution of the PC
fh_acs_scoring_by_county <- fh_acs_scoring_by_county %>% mutate(risk_score_all=rowSums(.[established_cols])) %>% mutate(risk_score_all=as.numeric(rescale(risk_score_all, to=c(0,100))))

plot(fh_acs_scoring_by_county$risk_score_pc_1, fh_acs_scoring_by_county$risk_score_all)
plot(fh_acs_scoring_by_county$risk_score_pc_2, fh_acs_scoring_by_county$risk_score_all)

fh_acs_scoring_by_county_distribute <- fh_acs_scoring_by_county %>% dplyr::select(c(established_cols, "county_code", "risk_score_both_pc", "risk_score_pc_1", "risk_score_pc_2", "risk_score_all", "total_population")) 

fh_acs_scoring_by_county_distribute <- fh_acs_scoring_by_county_distribute %>% mutate(risk_score = risk_score_both_pc)
fh_acs_scoring_by_county_distribute <- fh_acs_scoring_by_county_distribute %>% left_join(location_info)
fh_acs_scoring_by_county_distribute <- fh_acs_scoring_by_county_distribute %>% group_split(stateabbr) %>% map(function(.x) { .x %>% mutate(rank_in_state=rank(-risk_score)) }) %>% bind_rows()

DT::datatable(fh_acs_scoring_by_county_distribute %>% dplyr::select(county_code, stateabbr, risk_score, risk_score_pc_1, risk_score_all))
if(WRITE_OUTPUT) {
  write_rds(fh_acs_scoring_by_county, path = "fh_acs_covid_county_comm_score.rds")
  write_csv(fh_acs_scoring_by_county_distribute, path="fh_acs_covid_county_comm_score.csv")
}
```


# State COVID-19 Community Risk Score
```{r state risk score}
fh_acs_scoring_state <- fh_acs_scoring %>% dplyr::select(-c(pc_1, pc_2, risk_score_pc_1, risk_score_pc_2, risk_score_all, risk_score_both_pc, lat,long, geoid, fips_place_tract)) %>% mutate(total_population = total_males + total_females)

fh_acs_scoring_state <- fh_acs_scoring_state %>% cbind(fh_acs[non_na, c(established_disease_se, established_risk_se)])

cols_to_total <- c("male_over_65_pct","female_over_65_pct","avg_over_65_pct")
cols_total <- c("male_over_65_count","female_over_65_count","avg_over_65_count")

for(i in 1:length(cols_to_total)) {
  fh_acs_scoring_state <- fh_acs_scoring_state %>% mutate(!!cols_total[i] := .data[[cols_to_total[i]]]*total_population)
}
fh_acs_scoring_by_state <- fh_acs_scoring_state %>% group_by(stateabbr) %>% summarise_at(c(cols_total, 'total_population'), sum)

cols_to_prev <- c("male_over_65_pct","female_over_65_pct","avg_over_65_pct")
for(i in 1:length(cols_total)) {
  fh_acs_scoring_by_state <- fh_acs_scoring_by_state %>% mutate(!!cols_to_prev[i] := .data[[cols_total[i]]]/total_population)
}

weighted_state_prev_avgs <- function(d, prefix_cols) {
  avgs <- prefix_cols %>% map_dbl(~weighted_prev_avg(d, .x))
  place<- d[['stateabbr']][1]
  tibble(disease=prefix_cols, prevalence=avgs, stateabbr=place)
}

fh_acs_scoring_prevs_by_state <- fh_acs_scoring_state %>% group_by(stateabbr) %>% group_split(keep=TRUE) %>% map_df(~weighted_state_prev_avgs(.x, prefixes))
fh_acs_scoring_prevs_by_state$disease <- sprintf('%s_CrudePrev', fh_acs_scoring_prevs_by_state$disease)
fh_acs_scoring_prevs_by_state <- fh_acs_scoring_prevs_by_state %>% spread(disease, prevalence)

fh_acs_scoring_by_state <- fh_acs_scoring_by_state %>% left_join(fh_acs_scoring_prevs_by_state, by='stateabbr')

heatmap.2(cor(fh_acs_scoring_by_state[,established_cols]))

pc_by_state <- prcomp(fh_acs_scoring_by_state[,established_cols],center=T, scale.=T)
summary(pc_by_state)
pc_by_state$rotation


fh_acs_scoring_by_state$pc_1 <- as.numeric(pc_by_state$x[, 1])
fh_acs_scoring_by_state$pc_2 <- as.numeric(pc_by_state$x[, 2])

fh_acs_scoring_by_state <- fh_acs_scoring_by_state %>% mutate(risk_score_pc_1=as.numeric(rescale(pc_1,to=c(0,100))))
fh_acs_scoring_by_state <- fh_acs_scoring_by_state %>% mutate(risk_score_pc_2=as.numeric(rescale(-pc_2,to=c(0,100))))
fh_acs_scoring_by_state <- fh_acs_scoring_by_state %>% mutate(risk_score_both_pc=rescale(risk_score_pc_1*.61 + risk_score_pc_2*.24, to=c(0,100))) # rescale according the the contribution of the PC
fh_acs_scoring_by_state <- fh_acs_scoring_by_state %>% mutate(risk_score_all=rowSums(.[established_cols])) %>% mutate(risk_score_all=as.numeric(rescale(risk_score_all,to=c(0,100))))


fh_acs_scoring_by_state_distribute <- fh_acs_scoring_by_state %>% dplyr::select(c(established_cols, "stateabbr", "risk_score_both_pc", "risk_score_pc_1", "risk_score_pc_2", "risk_score_all", "total_population")) 

fh_acs_scoring_by_state_distribute <- fh_acs_scoring_by_state_distribute %>% mutate(rank_in_country=rank(risk_score_both_pc)) %>% mutate(risk_score = risk_score_both_pc)

DT::datatable(fh_acs_scoring_by_state_distribute %>% dplyr::select(stateabbr, risk_score,risk_score_all))
if(WRITE_OUTPUT) {
  write_rds(fh_acs_scoring_by_state, path = "fh_acs_covid_state_comm_score.rds")
  write_csv(fh_acs_scoring_by_state_distribute, path="fh_acs_covid_state_comm_score.csv")
}
```


